Interpolation and Extrapolation Performance Measurement of Analytical and ANN-Based Flow Laws for Hot Deformation Behavior of Medium Carbon Steel

### 2023-01-30 Article.tex
Final Submitted Article

\title{Interpolation and extrapolation-based performance measurement of Analytical and Artificial Neural Network models in predicting hot deformation behavior of a medium carbon steel}

@@

\begin{abstract}
\blfootnote{This article is an open access article distributed under the terms and
conditions of the Creative Commons Attribution (CC BY) license (\href{http://creativecommons.org/licenses/by/4.0}{http://creativecommons.org/licenses/by/4.0}).} In the present work, a critical analysis of the most commonly used analytical models and of recently introduced Artificial Neural Network based models is conducted to evaluate their prediction accuracy within and outside the experimental interval used to generate them.
To this end, the high-temperature deformation behavior of a medium carbon steel was studied over a wide range of strains ($0.0-0.7$), strain rates ($0.001~\ps-5~\ps$) and temperatures ($1050\celsius-1250\celsius$) using hot compression tests on a Gleeble-3800 thermomechanical simulator.
The experimental flow curves were then modeled using Johnson-Cook, Modified Zerilli-Armstrong, Hansel-Spittel, Arrhenius models, the recently proposed PTM model, as well as an Artificial Neural Network model with $3$ inputs, $2$ hidden layers of $15$ and $7$ neurons, respectively, and 1 output.
The Mean Absolute Relative Error ($\MARE$) and the Root Mean Square Error ($\RMSE$) values were used to quantify the predictive accuracy of the analyzed models.
The results indicated that the Johnson-Cook and Modified Zerilli-Armstrong models presented a significant error, $\MARE$ of $21.20\%$, while the Hansel-Spittel, PTM and Arrhenius models could predict the behavior of this alloy with an error of about $3.56\%$ error.
The Artificial Neural Network model, on the other hand, showed excellent agreement, with less than a $0.62\%$ error, between the predicted and experimental flow curves.
To further validate the performance of the above models, their ability to interpolate and extrapolate the experimental data was also tested.
The Hansel-Spittel, PTM and Arrhenius analytical models showed a good interpolation and extrapolation capability, with the notable superiority of the Arrhenius model over the other two.
The Artificial Neural Networks model was, however, still the most performant among all the models.
\end{abstract}

%%Graphical abstract
\begin{graphicalabstract}
%\includegraphics{grabs}
\end{graphicalabstract}

%%Research highlights
\begin{highlights}
\item Analytical and ANN based flow laws are identified from hot compression of medium carbon steel alloy.
\item Comparison of models is done using error measure with regard to the experiments.
\item The efficiency of each model is evaluated using interpolation and extrapolation techniques.
\item The Arrhenius model is the more accurate of all analytical models.
\item The ANN based model is the most accurate of all proposed models.
\end{highlights}

\begin{keyword}
Johnson-Cook \sep Zerilli-Armstrong \sep Arrhenius \sep Hansel-Spittel \sep Artificial Neural Network \sep interpolation \sep extrapolation
\end{keyword}

@@

Neural networks generally have two uses: classification and regression.
The first is the ability to classify data into different groups, and for example, to distinguish between images of cats and dogs.
The second corresponds to the universal approximation capacity of these neural networks, which is of interest to us herein, and thus, to the ability, after training, to predict the flow stress $\sigma^y$ values according to the input data, akin to what is done by the above identified analytical models.
The main difference is that this approximation is not linked to a fixed mathematical formulation (as in JC, MZA, AR, HS and PTM analytical models), but is only dependent on the data used for training, the number of layers, the number of neurons per layer, and the activation functions associated with the neurons of the network.
A feed-forward ANN, as used in our application, contains an input layer, an output layer and a number of hidden layers ($2$ in our case).
Each layer of neurons is connected to the one before it and the one after it by weighted connections.
Thus, all the neurons of the $k$ layer are connected to all the neurons of the ($k-1$) layer as shown in Figure \ref{fig:ANN-2HL}.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]
{Figures/ANN-scheme-2HL}
\caption{Multi-layer Artificial Neural Network architecture}
\label{fig:ANN-2HL}
\end{figure}

>>>>

\title{Interpolation and Extrapolation Performance Measurement of Analytical and ANN-Based Flow Laws for Hot Deformation Behavior of Medium Carbon Steel}

@@

\begin{abstract}
\blfootnote{This article is an open access article distributed under the terms and
conditions of the Creative Commons Attribution (CC BY) license (\href{http://creativecommons.org/licenses/by/4.0}{http://creativecommons.org/licenses/by/4.0}).} 
%In the present work, a critical analysis of the most commonly used analytical models and of recently introduced Artificial Neural Network based models is conducted to evaluate their prediction accuracy within and outside the experimental interval used to generate them.
%To this end, the high-temperature deformation behavior of a medium carbon steel was studied over a wide range of strains ($0.0-0.7$), strain rates ($0.001~\ps-5~\ps$) and temperatures ($1050\celsius-1250\celsius$) using hot compression tests on a Gleeble-3800 thermomechanical simulator.
%The experimental flow curves were then modeled using Johnson-Cook, Modified Zerilli-Armstrong, Hansel-Spittel, Arrhenius models, the recently proposed PTM model, as well as an Artificial Neural Network model with $3$ inputs, $2$ hidden layers of $15$ and $7$ neurons, respectively, and 1 output.
%The Mean Absolute Relative Error ($\MARE$) and the Root Mean Square Error ($\RMSE$) values were used to quantify the predictive accuracy of the analyzed models.
%The results indicated that the Johnson-Cook and Modified Zerilli-Armstrong models presented a significant error, $\MARE$ of $21.20\%$, while the Hansel-Spittel, PTM and Arrhenius models could predict the behavior of this alloy with an error of about $3.56\%$ error.
%The Artificial Neural Network model, on the other hand, showed excellent agreement, with less than a $0.62\%$ error, between the predicted and experimental flow curves.
%To further validate the performance of the above models, their ability to interpolate and extrapolate the experimental data was also tested.
%The Hansel-Spittel, PTM and Arrhenius analytical models showed a good interpolation and extrapolation capability, with the notable superiority of the Arrhenius model over the other two.
%The Artificial Neural Networks model was, however, still the most performant among all the models.
In the present work, a critical analysis of the most commonly used analytical models and recently introduced ANN-based models is performed to evaluate their predictive accuracy within and outside the experimental interval used to generate them.
The high temperature deformation behavior of a medium carbon steel was studied over a wide range of strains, strain rates, and temperatures using hot compression tests on a Gleeble-3800.
The experimental flow curves were modeled using Johnson-Cook, Modified-Zerilli-Armstrong, Hansel-Spittel, Arrhenius, and PTM models, as well as an ANN model.
The mean absolute relative error and root mean square error values were used to quantify the predictive accuracy of the models analyzed.
The results indicated that the Johnson-Cook and Modified-Zerilli-Armstrong models had a significant error, while the Hansel-Spittel, PTM, and Arrhenius models were able to predict the behavior of this alloy.
The ANN model showed excellent agreement, with less than a 0.62\% error, between the predicted and experimental flow curves.
To validate the performance, the ability to interpolate and extrapolate the experimental data was also tested.
The Hansel-Spittel, PTM, and Arrhenius models showed good interpolation and extrapolation capabilities.
However, the ANN model was the most powerful of all the models.
\end{abstract}

%%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics[width=\columnwidth]{Figures/GraphicalAbstract}
%\end{graphicalabstract}
%
%%%Research highlights
%\begin{highlights}
%\item Analytical and ANN based flow laws are identified from hot compression of medium carbon steel alloy.
%\item Comparison of models is done using error measure with regard to the experiments.
%\item The efficiency of each model is evaluated using interpolation and extrapolation techniques.
%\item The Arrhenius model is the more accurate of all analytical models.
%\item The ANN based model is the most accurate of all proposed models.
%\end{highlights}

\begin{keyword}
Artificial Neural Network\sep Constitutive flow Law\sep Analytical flow law\sep Interpolation\sep Extrapolation\sep Gleeble
\end{keyword}

@@

Neural networks generally have two uses: classification and regression.
The first is the ability to classify data into different groups, and for example, to distinguish between images of cats and dogs.
The second corresponds to the universal approximation capacity of these neural networks, which is of interest to us herein, and thus, to the ability, after training, to predict the flow stress $\sigma^y$ values according to the input data, akin to what is done by the above identified analytical models.
The main difference is that this approximation is not linked to a fixed mathematical formulation (as in JC, MZA, AR, HS and PTM analytical models), but is only dependent on the data used for training, the number of layers, the number of neurons per layer, and the activation functions associated with the neurons of the network.
A feed-forward ANN, as used in our application, contains an input layer, an output layer and a number of hidden layers ($2$ in our case).
Each layer of neurons is connected to the one before it and the one after it by weighted connections.
Thus, all the neurons of the $k$ layer are connected to all the neurons of the ($k-1$) layer as shown in Figure \ref{fig:ANN-2HL}.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]
{Figures/ANN-scheme-2HL}
\caption{Two hidden layers Artificial Neural Network architecture}
\label{fig:ANN-2HL}
\end{figure}
